Submodular Optimization for Control of Prosumer Networks

### 2016-04-19 Article.tex
corrections michel

We study the placement of storage in networks of prosumers that may both consume and produce electricity using renewable generators (DER). In this stochastic context, power imbalances are very likely, such that the system might lose synchrony and require control to bring it back to the synchronized state. We consider the control inputs as being the amount of power injected in the grid at the chosed storage locations. In a context where generators and loads are stochastic and not fixed, the optimal placement of the storages beforehand is not trivial. The dynamic is written as a system of differential equations and the physical constraints of the system are included. We then propose and validate an algorithm based on a submodular optimization that minimizes the average energy required to control the system.

@@

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Agents responsible for generation and load portfolio managment, the so-called prosumers \cite{Parag2016} because they may both produce and consume electricity, will provide services (generation, load shedding, frequency regulation) against remuneration. Insure grid stability within this uncertain context appears as a complex task.

@@

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. Re-balancing the production and the consumption is then necessary and requires that we have the capability to control the grid's dynamic. In \cite{Gkatzikis2015}, the authors study the placement of storage equipments within microgrids in a collaborative scheme. The authors solve an interesting optimization problem mixing placement-dimensioning of storage with utilization. Moreover, \cite{Gkatzikis2015} uses a Nash bargaining framework to determine how prosumers should share the costs and benefits of the storages. However, \cite{Gkatzikis2015} does not take into account the grid dynamic as well as its controllability. Control of smart grid systems has been recently studied in \cite{Farraj2015}, where the authors use linear-quadratic optimal control theory in order to compute DER outputs. Control inputs rely on information collected by sensors and phasor measurement units and a communication network is used for sharing the information. The performance is then studied under practical limitations such as latency, sampling rate, or signal-to-noise ratio. The objective of \cite{Farraj2015} is the control of the system when perturbations occur. The control signals are applied to the DER, meaning that \cite{Farraj2015} does not consider a system where storage would be present and used for control.

@@

The power grid dynamic is represented by a coupled oscillator network \cite{Filatrella2008}. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}.

@@

\section{Conclusion}
\label{sec:Conclusion}
In this paper, we considered the case of networks of prosumers which can behave as generators or loads depending on weather conditions. We believe that the originality of this work was to study the storage placement in such a stochastic context through the optimal control theory. This approach enabled us to take the grid's dynamic and its physical constraints into account. Because perturbations could be of any kind, and because we only place the storages once, we proposed an optimization method, that rely on submodular set functions and the gramian matrix, for minimizing the average energy that should be injected or absorbed by the storage equipments.

We believe that interesting work could be done by combining this model with real production and consumption data. There are indeed complex spatial and temporal correlations that impacts these distributions \cite{Gensollen2014}. Weather perturbations would therefore affect the nodes not completely at random, and affect how control should be designed.

>>>>

We propose here an optimization method for the placement of storage in networks of prosumers that may both consume and produce electricity using renewable generators (DER). In this stochastic context, power imbalances are very likely, such that the system might lose synchrony and require control to bring it back to the synchronized state. We consider the control inputs as being the amount of power injected in the grid at the chosed storage locations. In a context where generators and loads are stochastic and not fixed, the optimal placement of the storages beforehand is not trivial. We model the dynamic as a system of differential equations and the physical constraints of the system are included. We then propose and validate an algorithm based on a submodular optimization that minimizes the average energy required to control the system.

@@

The today centralized top-down architecture is set to evolve to more distributed and bi-directional systems. Furthermore, the emergence of renewable and stochastic distributed energy resources (DER) in the distribution networks will require more flexibility. Within these systems, it is assumed that multiple aggregation levels will be required in order to organize and optimize communication. Prosumer agents \cite{Parag2016} that both produce and consume electricity will provide services (generation, load shedding, frequency regulation) against remuneration. Insure grid stability within this uncertain context appears as a complex task.

@@

In this paper, we explore the use of storage in a network of prosumers whose power outputs are susceptible to change due to various external causes. In such a scenario with a high penetration of renewables, perturbations in the power distributions of the prosumers are very likely. Re-balancing the production and the consumption is then necessary and requires that we have the capability to control the grid's dynamic. In \cite{Gkatzikis2015}, the authors study the placement of storage equipments within microgrids in a collaborative scheme. The authors solve an interesting optimization problem mixing placement-dimensioning of storage with utilization. Moreover, \cite{Gkatzikis2015} uses a Nash bargaining framework to determine how prosumers should share the costs and benefits of the storages. However, \cite{Gkatzikis2015} does not take into account the grid dynamic as well as its controllability. Control of smart grid systems has been recently studied in \cite{Farraj2015}, where the authors use linear-quadratic optimal control theory in order to compute DER outputs. Control inputs rely on information collected by sensors and phasor measurement units and a communication network is used for sharing the information. The performance is then studied under practical limitations such as latency, sampling rate, or signal-to-noise ratio. The objective of \cite{Farraj2015} is the control of the system when perturbations occur. The control signals are used for adjusting the productions of the DER to some extent, but storage is not considered within this work.

@@

The power grid dynamic, where the prosumers powers vary in function of time, can be represented by a coupled oscillator network \cite{Filatrella2008}. Each node, representing a prosumer, is modeled as an oscillator with a phase angle and a natural frequency that depends on its production/consumption. The oscillators are interconnected such that their frequencies are impacted by the ones of their neighbors. As we will see, the phase angles dynamics can be written as a system of coupled second order differential equations \cite{Filatrella2008}.

@@

\section{Conclusion}
\label{sec:Conclusion}
In this paper, we considered the case of networks of prosumers which can behave as generators or loads depending on weather conditions. We proposed an optimization method of the storage placement in such a stochastic context through the optimal control theory. This approach enabled us to take the grid's dynamic and its physical constraints into account. Because perturbations could be of any kind and thus the control in an direction of the state space, our optimization method relies on submodular set functions and the gramian matrix, for minimizing the average energy that should be injected or absorbed by the storage equipments.

We believe that interesting work could be done by combining this model with real production and consumption data. There are indeed complex spatial and temporal correlations that impacts these distributions \cite{Gensollen2014}. In contrast, we considered that the control has to be in average in any direction of the state space but confronting this assumption with real traces could lead to interesting further research.

% Weather perturbations would therefore affect the nodes not completely at random, and affect how control should be designed.
